---
title: "script"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(jsonlite)

zip_fips <- read.csv("ZIP-COUNTY-FIPS.csv")

LIKERT_ORDERED <- c("Strongly disagree", "Disagree", "Neutral", "Agree", "Strongly agree")
responses <- fromJSON("../data/form_responses.json") %>% 
  separate(categories, sep = "; ", into = c("category1", "category2", "category3")) %>% 
  gather(category_num, category, category1:category3) %>% 
  filter(!is.na(category)) %>%
  mutate(
    eliminate_loans = factor(eliminate_loans, levels = LIKERT_ORDERED),
    income = factor(income, levels = c("Less than $40K", "$40K to $80K", "$80K to $125K", "$125K to $250K", "$250K to $350K", "$350K to $500K", "Greater than $500K")),
    net_price = factor(net_price, levels= c("0", "Less than $5K", "$10K to $15K", "$15K to $20K", "$20K to $25K", "$25K to $30K", "$30K to $35K", "$35K to $40K", "$45K to $50K", "$50K to $55K", "$55K to $60K", "$60K to $65K", "$65K to $70K", "$70K to $75K", "More than $75K")),
    affirmative_action = factor(affirmative_action, levels = LIKERT_ORDERED),
    political_af = factor(political_af, levels = c("Very conservative", "Somewhat conservative", "Somewhat liberal", "Very liberal", "I don't know", "Neither conservative nor liberal"))
  )

responses_ec <- responses %>% 
  separate(extracurriculars, sep = ", ", into = c("ec1", "ec2", "ec3", "ec4", "ec5", "ec6", "ec7", "ec8")) %>% 
  gather(extracurricular_num, extracurricular, ec1:ec8) %>% 
  filter(!is.na(extracurricular)) %>% 
  group_by(extracurricular) %>% 
  filter(n() > 10) %>% 
  ungroup()
```

## Preliminary looks

```{r}
# School by income
responses %>%
  ggplot(aes(income, fill = category)) +
  geom_bar(position = "fill") +
  coord_flip()

responses %>%
  ggplot(aes(school, fill = income)) +
  geom_bar(position = "fill") +
  coord_flip()

# Loans opinion by income
responses %>%
  ggplot(aes(income, fill = eliminate_loans)) +
  geom_bar(position = "fill") +
  scale_fill_manual(values = c("dodgerblue3", "dodgerblue", "gray", "firebrick1", "firebrick")) +
  coord_flip()
```

## Map data export shit

```{r}
responses_fips <- responses %>% 
  merge(zip_fips %>% select(ZIP, STCOUNTYFP) %>% rename(zipcode = ZIP), by = "zipcode") %>%
  rename(fips = STCOUNTYFP)

responses_fips %>% 
  group_by(fips) %>% 
  count() %>% 
  write.csv(file = "fips_counts.csv")  
```

## Cross sectional stuff

Students who had 504 plans both before and during high school are largely from public non-magnet, non-charter schools, but only half of students who had 504 plans only during high school are from public non-magnet, non-charter schools.
```{r}
responses %>%
  filter(plan_504 != "Never") %>%
  ggplot(aes(plan_504, fill = hs_type)) +
  geom_bar() +
  coord_flip()
```

Testing
```{r}
responses %>% 
  gather(key = test, value = test_count, sat_count, act_count) %>%  
  mutate(test_count = factor(test_count)) %>% 
  ggplot(aes(income, fill = test_count)) +
  geom_bar(position = "fill") +
  coord_flip()
```

```{r}
responses %>% 
  gather(key = test, value = test_count, sat_count, act_count) %>%  
  filter(test_count == 0) %>% 
  ggplot(aes(hs_type, fill = income)) +
  geom_bar() +
  coord_flip()
```

Legacy kids
* much more White kids than Blacks/Hispanics have legacy, and richer 
```{r}
responses %>%
  mutate(legacy = if_else(legacy_degree == "No", "No", "Yes")) %>% 
  ggplot(aes(legacy, fill = income)) +
  geom_bar()
```

Affirmative action
```{r}
responses %>%
  ggplot(aes(eliminate_loans, fill = net_price)) +
  geom_bar() +
  coord_flip()
```

Low-income students don't want Greek life.
```{r}
extracurriculars %>% 
  filter(category != "Native Hawaiian or Other Pacific Islander") %>% 
  ggplot(aes(category, fill = extracurricular)) +
  geom_bar(position = "fill") + 
  coord_flip()
  # facet_wrap(~ category)
```

## Extracurriculars

Conservatives slightly more likely to join political organizations and student government, liberals much more likely to join social justice/activism and culture or identity groups.
```{r}
responses_ec %>%
  filter(!political_af %in% c("I don't know", "Neither conservative nor liberal")) %>% 
  filter(!school %in% c("Barnard College", "General Studies")) %>%  # barnard not enough conservatives, gs not enough students
  filter(income %in% c("Greater than $500K", "$125K to $250K", "Less than $40K")) %>% # only enough conservatives in these buckets
  mutate(
    political_af = as.character(political_af),
    political_af = if_else(political_af == "Very conservative", "Somewhat conservative", political_af),
    political_af = if_else(political_af == "Very liberal", "Somewhat liberal", political_af),
  ) %>% 
  ggplot(aes(income, fill = extracurricular)) +
  geom_bar(position = "fill") + 
  coord_flip() +
  facet_grid(~ political_af, scales = "free")
```

Club sport participation rises with income
```{r}
responses_ec %>%
  # filter(extracurricular %in% c("Club/Intramural Sports")) %>% 
  ggplot(aes(income, group = extracurricular, color = extracurricular)) +
  geom_line(stat = "density") +
  coord_flip() +
  facet_grid(hs_type ~.)
```

Asian
- Richer = sports, visual and performing arts, 

Hispanic/Latino: Culture/identity interest goes down with in higher income brackets. 

```{r}
responses_ec %>%
  filter(!category %in% c("Native Hawaiian or Other Pacific Islander")) %>% 
  mutate(
    extracurricular = if_else(extracurricular %in% c("Culture or Identity Groups", "Political Organizations", "Social Justice/Activism"), extracurricular, "Other")
  ) %>% 
  ggplot(aes(income, fill = extracurricular)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~.)
```


## Socioeconomic diversity

Asians and Whites are spread out pretty evenly in all income brackets, but Blacks and Hispanics are skewed to low-income.

**More poor students are first generation.**
```{r}
responses %>% 
  filter(!category %in% c("Native Hawaiian or Other Pacific Islander")) %>% 
  mutate(
    category = factor(category, levels = c("White", "Asian", "Hispanic, Latino, or Spanish origin", "Black or African American", "American Indian or Alaskan Native", "Middle Eastern or North African"))
  ) %>% 
  ggplot(aes(income)) +
  geom_bar() +
  facet_grid(~ category) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
  )
```

Support for affirmative action among Asians and whites peaks don't show too strongly of a trend , but it seems that higher-income students tend to not support affirmative action. But
```{r}
responses %>% 
  group_by(income, category) %>% 
  filter(n() > 10) %>% 
  ungroup() %>% 
  ggplot(aes(income, fill = affirmative_action)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ .)
```


## Voting

```{r}
responses %>% 
  filter(vote_primary != "" && vote_gubernatorial != "" && vote_senate != "" && vote_house != "") %>% 
  filter(!political_af %in% c("I don't know", "Neither conservative nor liberal")) %>% 
  mutate(
    political_af = as.character(political_af),
    political_af = if_else(political_af == "Very conservative", "Somewhat conservative", political_af),
    political_af = if_else(political_af == "Very liberal", "Somewhat liberal", political_af),
  ) %>% 
  ggplot(aes(income, fill = vote_primary)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ .)
```

## College decisions factors

```{r}
decisions <- responses %>% 
  gather(decision_factor, decision_priority, starts_with("decision"))

decisions %>% 
  filter(category %in% c("White", "Hispanic, Latino, or Spanish origin", "Black or African American", "Asian")) %>%
  # filter(hs_type != "Homeschooled") %>% 
  ggplot(aes(income, fill = decision_priority)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ decision_factor)
```

## Columbia factors



```{r}
factors <- responses %>% 
  gather(factor, factor_priority, starts_with("factor_"))

factors %>% 
  filter(!factor %in% c("factor_cost", "factor_resources", "factor_core", "factor_abroad", "factor_research", "factor_extracurriculras", "factor_athletics")) %>% 
  filter(category %in% c("White", "Hispanic, Latino, or Spanish origin", "Black or African American", "Asian")) %>%
  mutate(
    factor_priority = if_else(factor_priority == "Strongly negative", "Negative", factor_priority),
    factor_priority = if_else(factor_priority == "Strongly positive", "Positive", factor_priority),
    factor_priority = factor(factor_priority, levels = c("Negative", "Indifferent", "Positive"))
  ) %>% 
  ggplot(aes(income, fill = factor_priority)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ factor)
```


```{r}
factors <- responses %>% 
  gather(factor, factor_priority, starts_with("factor_"))

factors %>% 
  filter(factor %in% c("factor_cost", "factor_resources",
                       #"factor_core", we have a question about this already
                       "factor_demographics",
                       "factor_greek",
                       #"factor_mental_health", we have a question about this already
                       "factor_social")) %>% 
  filter(category %in% c("White", "Hispanic, Latino, or Spanish origin", "Black or African American", "Asian")) %>%
  mutate(
    factor_priority = if_else(factor_priority == "Strongly negative", "Negative", factor_priority),
    factor_priority = if_else(factor_priority == "Strongly positive", "Positive", factor_priority),
    factor_priority = factor(factor_priority, levels = c("Negative", "Indifferent", "Positive"))
  ) %>% 
  ggplot(aes(income, fill = factor_priority)) +
  geom_bar(position = "fill") +
  coord_flip() +
  facet_grid(category ~ factor)
```


